Dynamics Analysis: The Sound of Teaching Music 2

### 2022-06-27 experiment-2/analysis/stats/dynamics.Rmd
update additional analysis (normalised per trial / participant)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

```{r vel-dyn-trial, echo = FALSE}
# For each individual
vel_dyn_trial <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
vel_dyn_trial
```

### Plot
#### Line graph
```{r vel-dyn-trial-line-forte, echo = FALSE, fig.height = 4}
ggline(dt_vel_dyn[Subcomponent == "Forte", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Forte") +
  scale_x_continuous(breaks = seq(1,8,1))
```

```{r vel-dyn-trial-line-piano, echo = FALSE, fig.height = 4}
ggline(dt_vel_dyn[Subcomponent == "Piano", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Piano") +
  scale_x_continuous(breaks = seq(1,8,1))
```

@@

```{r vel-diff-dyn-trial, echo = FALSE}
# For each individual
vel_diff_dyn_trial <- dt_vel_diff_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent,TrialNr)]
vel_diff_dyn_trial
```

### Plot
#### Line graph
```{r vel-diff-dyn-trial-line-ftop, echo = FALSE, fig.height = 4}
ggline(dt_vel_diff_dyn[Subcomponent == "FtoP", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Dynamics/Forte to Piano") +
  scale_x_continuous(breaks = seq(1,8,1))
```

```{r vel-diff-dyn-trial-line-ptof, echo = FALSE, fig.height = 4}
ggline(dt_vel_diff_dyn[Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Dynamics/Piano to Forte") +
  scale_x_continuous(breaks = seq(1,8,1))
```

@@

```{r vel-art, echo = FALSE}
# For each individual
vel_art <- dt_vel_art[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
vel_art
```

### Plot
#### Box plot
```{r vel-art-box,  echo = FALSE}
ggboxplot(dt_vel_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Articulation") +
  facet_grid(Subcomponent ~ .)
```

@@

```{r vel-art-trial, echo = FALSE}
# For each individual
vel_art_trial <- dt_vel_art[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
vel_art_trial
```

### Plot
#### Line graph
```{r vel-art-trial-line-legato, echo = FALSE, fig.height = 4}
ggline(dt_vel_art[Subcomponent == "Legato", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Articulation/Legato") +
  scale_x_continuous(breaks = seq(1,8,1))
```

```{r vel-art-trial-line-staccato, echo = FALSE, fig.height = 4}
ggline(dt_vel_art[Subcomponent == "Staccato", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Articulation/Staccato") +
  scale_x_continuous(breaks = seq(1,8,1))
```

@@

### Plot
#### Line graph
```{r vel-diff-artt-rial-line-ltos, echo = FALSE, fig.height = 4}
ggline(dt_vel_diff_art[Subcomponent == "LtoS", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Articulation/Legato to Staccato") +
  scale_x_continuous(breaks = seq(1,8,1))
```

```{r vel-diff-art-trial-line-stol, echo = FALSE, fig.height = 4}
ggline(dt_vel_diff_art[Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Articulation/Staccato to Legato") +
  scale_x_continuous(breaks = seq(1,8,1))
```

@@

```{r vel-diff-art-trial-all, echo = FALSE}
# Group mean
vel_diff_art_trial_all <- vel_diff_art_trial[Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
vel_diff_art_trial_all
```

@@

>>>>

```{r vel-dyn-trial, echo = FALSE}
# For each individual
vel_dyn_trial
```

### Plot
#### Line graph
```{r vel-dyn-trial-line-forte, echo = FALSE, fig.height = 4}
ggline(vel_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Forte") +
  scale_x_continuous(breaks = seq(1,8,1))
```

```{r vel-dyn-trial-line-piano, echo = FALSE, fig.height = 4}
ggline(vel_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Piano") +
  scale_x_continuous(breaks = seq(1,8,1))
```

@@

```{r vel-diff-dyn-trial, echo = FALSE}
# For each individual
vel_diff_dyn_trial
```

### Plot
#### Line graph
```{r vel-diff-dyn-trial-line-ftop, echo = FALSE, fig.height = 4}
ggline(vel_diff_dyn_trial[Subcomponent == "FtoP"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Dynamics/Forte to Piano") +
  scale_x_continuous(breaks = seq(1,8,1))
```

```{r vel-diff-dyn-trial-line-ptof, echo = FALSE, fig.height = 4}
ggline(vel_diff_dyn_trial[Subcomponent == "PtoF"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Dynamics/Piano to Forte") +
  scale_x_continuous(breaks = seq(1,8,1))
```

@@

```{r vel-art, echo = FALSE}
# For each individual
vel_art_trial <- dt_vel_art[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]

vel_art <- vel_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
vel_art
```

### Plot
#### Box plot
```{r vel-art-box,  echo = FALSE}
ggboxplot(vel_art_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Articulation") +
  facet_grid(Subcomponent ~ .)
```

@@

```{r vel-art-trial, echo = FALSE}
# For each individual
vel_art_trial
```

### Plot
#### Line graph
```{r vel-art-trial-line-legato, echo = FALSE, fig.height = 4}
ggline(vel_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Articulation/Legato") +
  scale_x_continuous(breaks = seq(1,8,1))
```

```{r vel-art-trial-line-staccato, echo = FALSE, fig.height = 4}
ggline(vel_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Articulation/Staccato") +
  scale_x_continuous(breaks = seq(1,8,1))
```

@@

### Plot
#### Line graph
```{r vel-diff-artt-rial-line-ltos, echo = FALSE, fig.height = 4}
ggline(vel_diff_art_trial[Subcomponent == "LtoS"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Articulation/Legato to Staccato") +
  scale_x_continuous(breaks = seq(1,8,1))
```

```{r vel-diff-art-trial-line-stol, echo = FALSE, fig.height = 4}
ggline(vel_diff_art_trial[Subcomponent == "StoL"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Articulation/Staccato to Legato") +
  scale_x_continuous(breaks = seq(1,8,1))
```

@@

```{r vel-diff-art-trial-all, echo = FALSE}
# Group mean
vel_diff_art_trial_all <- vel_diff_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
vel_diff_art_trial_all
```

@@

# Additional analysis{#additional}
- Normalise KV difference by the range of KV value between forte and piano for each participant.

## Normalised KV difference per trial

```{r, echo = FALSE, message = FALSE}
# normalise diff by KV range
vel_diff_dyn_trial$NormalisedDiff <- 0
for (i in unique(vel_diff_dyn_trial$SubNr)){
  for (m in unique(vel_diff_dyn_trial[SubNr == i]$TrialNr)){
    # teaching
    # FtoP
    vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching" & Subcomponent == "FtoP"]$NormalisedDiff <- vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching" & Subcomponent == "FtoP"]$Mean/(max(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean)-min(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean))
    # PtoF
    vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching" & Subcomponent == "PtoF"]$NormalisedDiff <- vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching" & Subcomponent == "PtoF"]$Mean/(max(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean)-min(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean))

    # performing
    # FtoP
    vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing" & Subcomponent == "FtoP"]$NormalisedDiff <- vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing" & Subcomponent == "FtoP"]$Mean/(max(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean)-min(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean))
    # PtoF
    vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing" & Subcomponent == "PtoF"]$NormalisedDiff <- vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing" & Subcomponent == "PtoF"]$Mean/(max(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean)-min(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean))

    # error messages for non-existing trials but it's not reflected on vel_diff_dyn_trial anyway
  }
}

# each participant
vel_diff_dyn_normalised <- vel_diff_dyn_trial[, .(N = .N, NormalisedMean = mean(NormalisedDiff), SD = sd(NormalisedDiff), SEM = sd(NormalisedDiff)/sqrt(.N)), by = .(SubNr, Condition, Skill, Subcomponent)]

# group
vel_diff_dyn_normalised_all <- vel_diff_dyn_normalised[, .(N = .N, Mean = mean(NormalisedMean), SD = sd(NormalisedMean), SEM = sd(NormalisedMean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
vel_diff_dyn_normalised_all
```

### Plot
```{r, echo = FALSE}
vel_diff_dyn_plot_2 <- data.table(subject = vel_diff_dyn_normalised[Condition == "teaching"]$SubNr, teaching = vel_diff_dyn_normalised[Condition == "teaching"]$NormalisedMean, performing = vel_diff_dyn_normalised[Condition == "performing"]$NormalisedMean, subcomponent = vel_diff_dyn_normalised[Condition == "teaching"]$Subcomponent)

ggpaired(vel_diff_dyn_plot_2, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised KV (0-127)", title = "KV-Diff: Dynamics")
```

### Stats
```{r, echo = FALSE}
# stats
vel_diff_dyn_aov_3 <- aov_car(NormalisedMean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_dyn_normalised)
summary(vel_diff_dyn_aov_3)
```

## Normalised KV difference per participant
```{r, echo = FALSE}
vel_diff_dyn$NormalisedDiff <- 0
for (i in unique(vel_diff_dyn_trial$SubNr)){
  # teaching
  # FtoP
  vel_diff_dyn[SubNr == i & Condition == "teaching" & Subcomponent == "FtoP"]$NormalisedDiff <- vel_diff_dyn[SubNr == i & Condition == "teaching" & Subcomponent == "FtoP"]$Mean/(max(vel_dyn[SubNr == i & Condition == "teaching"]$Mean)-min(vel_dyn[SubNr == i & Condition == "teaching"]$Mean))
  # PtoF
  vel_diff_dyn[SubNr == i & Condition == "teaching" & Subcomponent == "PtoF"]$NormalisedDiff <- vel_diff_dyn[SubNr == i & Condition == "teaching" & Subcomponent == "PtoF"]$Mean/(max(vel_dyn[SubNr == i & Condition == "teaching"]$Mean)-min(vel_dyn[SubNr == i & Condition == "teaching"]$Mean))
  # performing
  # FtoP
  vel_diff_dyn[SubNr == i & Condition == "performing" & Subcomponent == "FtoP"]$NormalisedDiff <- vel_diff_dyn[SubNr == i & Condition == "performing" & Subcomponent == "FtoP"]$Mean/(max(vel_dyn[SubNr == i & Condition == "performing"]$Mean)-min(vel_dyn[SubNr == i & Condition == "performing"]$Mean))
  # PtoF
  vel_diff_dyn[SubNr == i & Condition == "performing" & Subcomponent == "PtoF"]$NormalisedDiff <- vel_diff_dyn[SubNr == i & Condition == "performing" & Subcomponent == "PtoF"]$Mean/(max(vel_dyn[SubNr == i & Condition == "performing"]$Mean)-min(vel_dyn[SubNr == i & Condition == "performing"]$Mean))
}

# group
vel_diff_dyn_normalised_all_2 <- vel_diff_dyn[, .(N = .N, Mean = mean(NormalisedDiff), SD = sd(NormalisedDiff), SEM = sd(NormalisedDiff)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
vel_diff_dyn_normalised_all_2
```

### Plot
```{r, echo = FALSE}
vel_diff_dyn_plot_3 <- data.table(subject = vel_diff_dyn[Condition == "teaching"]$SubNr, teaching = vel_diff_dyn[Condition == "teaching"]$NormalisedDiff, performing = vel_diff_dyn[Condition == "performing"]$NormalisedDiff, subcomponent = vel_diff_dyn[Condition == "teaching"]$Subcomponent)

ggpaired(vel_diff_dyn_plot_3, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised KV (0-127)", title = "KV-Diff: Dynamics")
```

### Stats
```{r, echo = FALSE}
# stats
vel_diff_dyn_aov_4 <- aov_car(NormalisedDiff ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_dyn)
summary(vel_diff_dyn_aov_4)
```
